---
title: "Advanced Plots in R with str_glue()"
author: "Jason Dexter"
date: '2020-05-03'
slug: advanced-plots-in-r-with-str_glue
categories: []
tags:
- R
- stringr
- ggplot
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    fig_height: 3
    fig_width: 8
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```


### Let's Dive Into an Example

I'd like to share a technique using `str_glue()` that I learned from [Matt Dancho](https://www.linkedin.com/in/mattdancho/){target="_blank"}, a Data-Science instructor at Business Science University. Check out my favorite course here: [Business Analysis With R](http://bit.ly/2Z9SHsd){target="_blank"}.

`str_glue()` from the `stringr` library is one of my favorite functions in `R` - It's super helpful for wrangling and manipulating text in preparation for building plots. 

### Let's Get Some Data 🤓

The [Tidy Tuesday Project](https://github.com/rfordatascience/tidytuesday){target="_blank"} is an awesome repository of useful data for practicing your data `Wrangling` skills.

We will work with the 'San Francisco Trees' data as a case study for using `str_glue()` for advanced plotting techniques.

```{r}
library(tidyverse)
library(stringr)
library(tidyquant)
library(scales)
library(DataExplorer)

sf_trees_data_raw_tbl <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')
```

### Data Exploration 

Let's take a quick peak and inspect the SF Trees Data.

```{r, fig.height=4}
plot_missing(sf_trees_data_raw_tbl, ggtheme = theme_tq())
```

This is a fairly small data-set with only 12 columns. For the purpose of demonstration, let's see what we can do with just the species column.

The Coastal Redwoods in the SF area are incredible and one of my favorite species. I'm curious if other species of Redwoods are in SF and if so, at what proportions do they exist relative to the Coastal Redwoods. 

### Data Wrangling

```{r}
# Data Wrangling
redwood_tbl <- sf_trees_data_raw_tbl %>% 
    
    # select species and filter to redwood only
    select(species) %>% 
    filter(str_detect(species, pattern = 'Redwood')) %>% 
    
    # break up species and common-name into separate columns
    separate(
        col = species,
        into = c("species", "common_name"),
        sep  = ' :: ',
        remove = FALSE) %>% 
    
    # calculate absolute and relative distributions
    count(species, common_name) %>% 
    mutate(pct = n / sum(n),
           pct_text = percent(pct)) %>% 
    arrange(desc(pct))
```

### Summary Table

Let's take a look at our findings.


```{r}
rmarkdown::paged_table(redwood_tbl %>% select(-pct))
```


As I expected, the Coastal Redwoods are the dominant species in San Francisco.

And while the table is good, lets craft an awesome plot to display these results.

### The Power of `str_glue()`

With a couple lines of `code` we can build our label for plotting. As you can see, we can add `arguments` directly from our table using curly brackets `{}` - honestly, the options are endless for how creative you can get with `stringing` together text and adding labels to your plots. 

```{r}
# Data Wrangling
redwood_labeled_text_tbl <- redwood_tbl %>% 
    
    # label text
    mutate(label_text = str_glue(
        'Scientific Name:
        {species}
        Count: {n} of {sum(n)}
        Pct (%): {pct_text}'
    ),
    
    # add 'forward-slash' to wrap-text on our plots y-axis
    common_name = str_replace(common_name, pattern = ' ', replacement = ' \n ')) %>% 
    
    # reorder factors based on percent rank
    mutate(common_name = common_name %>% fct_reorder(pct))
```


### Manipulated Text (ready to plot)

Here is the manipulated text that will be useful once we plot these data; this setup is critical for the labels on our final plot e.g., the `\n` will wrap the text at those locations.

```{r}
redwood_labeled_text_tbl %>% select(label_text) %>% rmarkdown::paged_table()
```


### Data Visualization

Now that we've done our `Data Wrangling`, lets get into a bit of `Data Visualization`.

```{r fig.height=2.8, fig.width=8}
# Save Plot
g <- redwood_labeled_text_tbl %>% 
    
    ggplot(aes(pct, common_name), color = '#2c3e50') +
    
    # Geometries
    geom_segment(aes(xend = 0, yend = common_name), size = 2) +
    geom_point(aes(size = 5)) +
    geom_label(aes(label = label_text),hjust = "inward",size = 3) +

    # Formatting
    scale_x_continuous(labels = scales::percent_format()) +
    labs(title = str_glue("Redwoods of San Francisco (SF)"),
        subtitle = str_glue(
          "As expected, the coastal Redwoods make up the largest proportion.
          Dawn Redwoods were once thought to be extinct, so low #'s aren't suprising.
          Siera Redwoods grow at high elev. and low numbers are expected"),
        caption = str_glue("The Coastal Redwood is the dominant species in SF."),
        x = "", y = "") +
    theme_tq() + theme(legend.position = "none",
          plot.title = element_text(face = "bold"))
```

## Display Awesome Plot 

And here is our plot with the engineered labels from the last few steps. And that's just one example of why I love `str_glue()` - Simply Awesome!

```{r echo=FALSE, fig.height=4.5, fig.width=8}
# Show Plot
g
```

That's it for today! 

We used `str_glue()` to manipulate our text and add awesome labels to our `ggplot()` - the plot is now `Business-Ready` 👍

Stay tuned for more posts on Wrangling + Exploring data (`programatically`).


### Stay Connected

Enter your [Email Here](https://tinyletter.com/dexters-analytics){target="_blank"} to receive notifications when new content is posted.<br/>  

PS: Be Kind and Tidy your Data 😎<br/>
<br/>


### Learn R Fast 👇

Interested in expediting your learning path? Head on over to [Business Science](https://www.business-science.io/?affcode=173166_vnvxtqbd){target="_blank"} and join me on the journey.

[Business Science: FREE Jumpstart Data-Science Course](https://university.business-science.io/p/jumpstart-with-r/?affcode=173166_vnvxtqbd){target="_blank"} (opened for a limited time)





```{r include=FALSE}
knitr::knit_exit()
```


### test?
